---
title: "p8105_hw1_zc2822"
author: Zhengyong Chen
date: "2024-09-19"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

# Problem 1

## Load the penguins dataset.

```{r}
data("penguins", package = "palmerpenguins")
```

## Describe the penguins dataset.

```{r}
variable_names = names(penguins)
species_names = levels(penguins$species)
island_names = levels(penguins$island)
bill_length_head10 = head(penguins$bill_length_mm, 10)
bill_depth_head10 = head(penguins$bill_depth_mm, 10)
flipper_length_head10 = head(penguins$flipper_length_mm, 10)
body_mass_head10 = head(penguins$body_mass_g, 10)

list("variable names" = variable_names,
     "species names" = species_names,
     "island names" = island_names,
     "head 10 of bill length" = bill_length_head10,
     "head 10 of bill depth" = bill_depth_head10,
     "head 10 of flipper length" = flipper_length_head10,
     "head 10 of body mass" = body_mass_head10)
```


```{r size}
list("number of rows" = nrow(penguins),
     "number of columns" = ncol(penguins))
```

```{r}
mean_flipper_length = mean(penguins$flipper_length_mm, na.rm = TRUE)
cat("mean flipper length:", mean_flipper_length)
```


## Using inline R code:

The data in the penguins dataset contains information about 8 variables: `r colnames(penguins)`.

The name of species: `r levels(penguins$species)`.

The name of islands: `r levels(penguins$island)`.

The data contain `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.


## Scatterplot

```{r scatterplot, warning = FALSE}
scatter_plot = ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
                      geom_point() +
                      labs(x = "Bill Length (mm)", y = "Flipper Length (mm)")
                      
print(scatter_plot)

ggsave("scatterplot_flipper_vs_bill.png", plot = scatter_plot)
```

# Problem 2

## Create the data frame.

```{r}
set.seed(1234)

df = tibble(
  random_sample = rnorm(10),
  logical_vector = random_sample > 0,
  character_vector = letters[1:10],
  factor_vector = factor(rep(c("low", "medium", "high"), length.out = 10))
)

print(df)
```

Try to calculate mean:

```{r}
df |> pull(random_sample) |> mean()
df |> pull(logical_vector) |> mean()
df |> pull(character_vector) |> mean()
df |> pull(factor_vector) |> mean()
```

I can only calculate the mean of random_sample and logical_vector.

```{r, eval = FALSE}
df |> pull(logical_vector) |> as.numeric()
df |> pull(character_vector) |> as.numeric()
df |> pull(factor_vector) |> as.numeric()
```

When applying as.numeric():

1. In the logical vector, TRUE was converted to 1, FALSE was converted to 0, because logical variables can be seen as numeric variables;

2. All characters in the character vector turned into NA, because the characters can not be converted into numeric ones directly;

3. In the factor vector, different level was converted into different integers, because factors are stored as integers.

For the logical vector, since TRUE was converted to 1, FALSE was converted to 0, so the mean can be calculated directly;

Character variables can not be directly converted to numeric, so calculating their mean results in an error;

Factor variables may have a mean after converted to numeric, but it does not have a meaningful explanation.


